# Point Cloud Processing Results

## Processing Time

let's look into the processing time comparison across data sets

the `cloud2trees::cloud2trees()` program automatically tracks processing time for all sections and stores the tracking data in the "point_cloud_processing_delivery" directory in a file called "processed_tracking_data.csv"

```{r}
dir_temp <- "../data"
# what processing data?
df_temp <-
  list.files(
    dir_temp
    , pattern = "processed_tracking_data.csv"
    , recursive = T
  ) %>% 
  dplyr::tibble() %>% 
  setNames("fpath") %>% 
  dplyr::mutate(
    data_desc = stringr::word(fpath, 2, sep = "/") %>% 
      stringr::str_remove_all("_processing") %>% 
      stringr::str_replace_all("_"," ") %>% 
      stringr::str_squish() %>% 
      toupper()
    , data_type = data_desc %>% 
      stringr::str_remove_all("[0-9]") %>% 
      stringr::str_squish()
    , study_site = stringr::word(fpath, sep = "/")
    , fpath = file.path(dir_temp, fpath)
    , fdir = dirname(fpath)
  )
# read in processing data
df_temp <- 1:nrow(df_temp) %>% 
  purrr::map(\(x)
    readr::read_csv(
      df_temp$fpath[x]
      , show_col_types = F
      , progress = F
    ) %>%
    dplyr::mutate(
      study_site = df_temp$study_site[x]
      , data_desc = df_temp$data_desc[x]
    )
  ) %>% 
  dplyr::bind_rows() %>% 
  dplyr::inner_join(df_temp, by = dplyr::join_by(study_site,data_desc))
# add in study bounds
study_sites_processing_sf <- 
  study_sites_sf %>% 
  dplyr::select(study_site, study_site_lab) %>% 
  dplyr::inner_join(
    df_temp
    , by = "study_site"
    , relationship = "one-to-many"
  )
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

let's relativize and proportionalize the tracking data

```{r}
# aggregate the total processing time
study_sites_processing_sf <- 
  study_sites_processing_sf %>% 
  dplyr::mutate(
    timer_total_time_mins = timer_cloud2raster_mins + timer_raster2trees_mins +
      timer_trees_dbh_mins + timer_trees_cbh_mins + timer_trees_type_mins +
      timer_trees_hmd_mins + timer_trees_biomass_mins + timer_write_data_mins
    , timer_tree_extraction_mins = timer_cloud2raster_mins + timer_raster2trees_mins
    , las_area_ha = (las_area_m2/10000)
    , points_m2 = number_of_points/las_area_m2
    # relative
    , dplyr::across(
      .cols = tidyselect::starts_with("timer_") & tidyselect::ends_with("_mins")
      , .fns = ~ (.x*60)/las_area_ha # sec/ha
      # , .fns = ~ .x/las_area_ha # min/ha
      , .names = "{.col}_secperha"
      # , .names = "{.col}_minperha"
    )
    # proportion
    , dplyr::across(
       .cols = c(timer_tree_extraction_mins,
        timer_trees_dbh_mins, timer_trees_cbh_mins, timer_trees_type_mins,
        timer_trees_hmd_mins, timer_trees_biomass_mins, timer_write_data_mins)
       , .fns = ~ .x/timer_total_time_mins
       , .names = "{.col}_pct"
    )
  )
study_sites_processing_sf %>% dplyr::glimpse()
```

that's a lot of tracking data, let's look at the main timing parameters

```{r}
# format data for plotting/tabling
table_temp <-
  study_sites_processing_sf %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(
    study_site, study_site_lab, data_desc, data_type, number_of_points, las_area_ha, points_m2
    , c(timer_tree_extraction_mins,
      timer_trees_dbh_mins, timer_trees_cbh_mins, timer_trees_type_mins,
      timer_trees_hmd_mins, timer_trees_biomass_mins, timer_write_data_mins
      , timer_total_time_mins
      , c(tidyselect::ends_with("_pct") & tidyselect::starts_with("timer_"))
      , c(tidyselect::ends_with("_secperha") & tidyselect::starts_with("timer_"))
    )
  ) %>% 
  tidyr::pivot_longer(
    cols = -c(study_site, study_site_lab, data_desc, data_type, number_of_points, las_area_ha, points_m2)
  ) %>% 
  dplyr::mutate(
    units = stringr::word(name, -1, sep = "_")
    , section = name %>% 
      stringr::str_remove_all("timer_") %>% 
      stringr::str_remove_all("_mins") %>% 
      stringr::str_remove_all("_pct") %>% 
      stringr::str_remove_all("_secperha")
  ) %>% 
  dplyr::select(-name) %>% 
  # dplyr::count(units)
  tidyr::pivot_wider(names_from = units, values_from = value) %>% 
  dplyr::mutate(
    mins_lab = scales::comma(mins,accuracy = 0.1)
    , perha_lab = scales::comma(secperha,accuracy = 0.01)
    , pct_lab = scales::percent(pct,accuracy = 0.1)
    # site lab
    , big_lab = stringr::str_c(
      data_desc
      , paste0("area: ", scales::comma(las_area_ha, accuracy = 1, suffix = " ha"))
      , paste0("points: ", scales::comma(number_of_points, accuracy = 0.1, scale = 1/1000000, suffix = "M"))
      , paste0("points m<sup>-2</sup>: ", scales::comma(points_m2, accuracy = 0.1))
      , sep = "<br>"
    )
    , big_lab_ggplot = stringr::str_c(
      data_desc
      , paste0("area: ", scales::comma(las_area_ha, accuracy = 1, suffix = " ha"))
      , paste0("points: ", scales::comma(number_of_points, accuracy = 0.1, scale = 1/1000000, suffix = "M"))
      , paste0("points/m2: ", scales::comma(points_m2, accuracy = 0.1))
      , sep = "\n"
    )
  ) %>%  
  # dplyr::count(section)
  dplyr::filter(!is.na(mins)) %>% 
  # section lab after filter so factor doesn't have rand levels
  dplyr::mutate(
    section = section %>% 
      stringr::str_replace_all("_", " ") %>% 
      stringr::str_remove_all("time") %>% 
      stringr::str_replace_all("dbh", "DBH") %>% 
      stringr::str_replace_all("cbh", "CBH") %>% 
      stringr::str_replace_all("hmd", "HMD") %>% 
      stringr::str_squish() %>% 
      forcats::fct_inorder()
  )
  
# table it
table_temp %>% 
  dplyr::select(study_site_lab,big_lab,section,mins_lab, perha_lab, pct_lab) %>% 
  kableExtra::kbl(
    caption = "Point cloud processing section run time"
    , col.names = c(
      "Site", "Data"
      , "Processing section"
      , "time (minutes)"
      , "seconds per ha"
      , "% of total time"
      )
    , escape = F
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::collapse_rows(columns = 1:2, valign = "top")

```

that's a lot of numbers to digest, let's plot the data

### Point Cloud Processing Time versus Point Density

```{r}
# per ha time based on point density
ggplot2::ggplot(
  data = table_temp %>% dplyr::filter(section == "total")
  , mapping = ggplot2::aes(y = secperha, x = points_m2, color = data_type)
) +
  ggplot2::geom_point(size = 4, alpha = 0.9) +
  ggplot2::scale_color_viridis_d(option = "magma", begin = 0.1, end = 0.5) +
  ggplot2::labs(
    x = latex2exp::TeX("points $m^{-2}$")
    , y = latex2exp::TeX("total seconds $ha^{-1}$")
    , color = ""
    , subtitle = "Point Cloud Processing Time versus Point Density"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "top")
```

### Point Cloud Processing Time by Section (%)

```{r}
table_temp %>% 
  dplyr::filter(section != "total") %>% 
  dplyr::mutate(section = forcats::fct_rev(section)) %>% 
ggplot2::ggplot(
  mapping = ggplot2::aes(y = big_lab_ggplot, x = pct, fill = section, group = section)
) +
  ggplot2::geom_col(
    width = 0.7, alpha=0.8
  ) +
  ggplot2::geom_text(
    mapping = ggplot2::aes(
        label = scales::percent(ifelse(pct>=0.06,pct,NA), accuracy = 1)
        , fontface = "bold"
      )
    , position = ggplot2::position_stack(vjust = 0.5)
    , color = "black", size = 4
  ) +
  ggplot2::facet_wrap(facets = dplyr::vars(study_site_lab), scales = "free_y") + 
  ggplot2::scale_fill_viridis_d(option = "turbo", begin = 0.1, end = 0.9) +
  ggplot2::scale_x_continuous(labels = scales::percent_format()) +
  labs(
    fill = "", y = ""
    , x = "% Point Cloud Total Processing Time"
    , subtitle = "Point Cloud Processing Time by Section"
  ) +
  theme_light() +
  theme(
    legend.position = "top"
    , legend.direction  = "horizontal"
    , legend.title = element_text(size=7)
    , axis.title.x = element_text(size=10, face = "bold")
    , axis.title.y = element_text(size = 8)
    , axis.text.x = element_blank()
    , axis.text.y = element_text(color = "black",size=10, face = "bold")
    , axis.ticks.x = element_blank()
  ) +
  guides(
    fill = guide_legend(nrow = 3, byrow = T, reverse = T, override.aes = list(alpha = 0.9))
  )
```

### Point Cloud Processing Time by Section (total)

```{r}
 table_temp %>% 
   dplyr::filter(section != "total") %>% 
   dplyr::mutate(section = forcats::fct_rev(section)) %>% 
 ggplot2::ggplot(
   mapping = ggplot2::aes(y = big_lab_ggplot, x = secperha, fill = section, group = section)
 ) +
  geom_text(
    data = table_temp %>% dplyr::filter(section == "total")
    , mapping = ggplot2::aes(
      y = big_lab_ggplot
      , x = secperha
      , label = scales::comma(secperha,accuracy=0.1,suffix = "\ntotal")
      , fontface = "bold"
    )
    , color = "black", size = 2.3
    , hjust = -0.1
  ) +
   ggplot2::geom_col(
     width = 0.7, alpha=0.8
   ) +
   ggplot2::geom_text(
     mapping = ggplot2::aes(
         label = scales::comma(ifelse(secperha>=7.5,secperha,NA), accuracy = 0.1)
         , fontface = "bold"
       )
     , position = ggplot2::position_stack(vjust = 0.5)
     , color = "black", size = 3
   ) +
   ggplot2::facet_wrap(facets = dplyr::vars(study_site_lab), scales = "free_y") + 
   ggplot2::scale_fill_viridis_d(option = "turbo", begin = 0.1, end = 0.9) +
   ggplot2::scale_x_continuous(labels = scales::comma_format(), expand = ggplot2::expansion(mult = c(0,0.1))) +
   labs(
     fill = "", y = ""
     , x = latex2exp::TeX("seconds $ha^{-1}$")
     , subtitle = "Point Cloud Processing Time by Section"
   ) +
   theme_light() +
   theme(
     legend.position = "top"
     , legend.direction  = "horizontal"
     , legend.title = element_text(size=7)
     , axis.title.x = element_text(size=10, face = "bold")
     , axis.title.y = element_text(size = 8)
     , axis.text.x = element_blank()
     , axis.text.y = element_text(color = "black",size=10, face = "bold")
     , axis.ticks.x = element_blank()
   ) +
   guides(
     fill = guide_legend(nrow = 3, byrow = T, reverse = T, override.aes = list(alpha = 0.9))
   )

```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Silvicultural Metrics

`cloud2trees` has some internal functions to identify `cloud2trees::cloud2trees()` outputs, read the outputs, and crop them to a given AOI, however, these are internal to the program (for now). We can use `cloud2trees::cloud2trees_to_lanl_trees()` (detailed in the [next section](#lanl_trees)) to access this functionality since all we need to do is clip the tree list which we can then aggregate to get common silvicultural metrics

### Example

here is an example to show the data we'll get back from the `calc_silv_metrics()` function we define below

```{r, results=F, message=FALSE}
# here's an example
# study_sites_processing_sf %>% dplyr::slice(6) %>% dplyr::glimpse()
# using the internal fn's would be: read_trees_flist() %>% clip_tree_list_aoi()
ans_temp <- cloud2trees::cloud2trees_to_lanl_trees(
  input_dir = study_sites_processing_sf$fdir[6]
  , study_boundary = study_sites_processing_sf %>% dplyr::slice(6)
  , bbox_aoi = F, buffer = 0 # this is the key to get only trees in our AOI
)
# ans_temp %>% names()
```

let's make a function to summarize the data and create common silvicultural metrics within our stand boundary

```{r}
###################################################################################
# define a function to convert columns in data frame from metric to imperial
# see: 
# https://www.forestnb.com/archives/forest-nb-news/resources/conversions/
# https://www.ars.usda.gov/is/np/agbyproducts/agbyappendix.pdf
###################################################################################
calc_imperial_units_fn <- function(df) {
  df %>% 
  # convert to imperial units
    dplyr::mutate(
      dplyr::across(
        .cols = tidyselect::ends_with("_cm")
        , ~ .x * 0.394
        , .names = "{.col}_in"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m")
        , ~ .x * 3.281
        , .names = "{.col}_ft"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m2_per_ha")
        , ~ .x * 4.359
        , .names = "{.col}_ftac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_kg_per_ha")
        , ~ .x * 0.892178
        , .names = "{.col}_lbsac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_per_ha") & 
          !tidyselect::ends_with("_m2_per_ha") & 
          !tidyselect::ends_with("_kg_per_ha")
        , ~ .x * 0.405
        , .names = "{.col}_ac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_area_ha")
        , ~ .x * 2.471
        , .names = "{.col}_ac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("kg_per_m2")
        , ~ .x * 0.20482
        , .names = "{.col}_lbsft2"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("kg_per_m3")
        , ~ .x * 0.0624
        , .names = "{.col}_lbsft3"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m2") & !tidyselect::ends_with("per_m2")
        , ~ .x * 10.764
        , .names = "{.col}_ft2"
      )
    ) %>%
    dplyr::rename_with(
      .fn = function(x){dplyr::case_when(
        stringr::str_ends(x,"_cm_in") ~ stringr::str_replace(x,"_cm_in","_in")
        , stringr::str_ends(x,"_m_ft") ~ stringr::str_replace(x,"_m_ft","_ft")
        , stringr::str_ends(x,"_m2_per_ha_ftac") ~ stringr::str_replace(x,"_m2_per_ha_ftac","_ft2_per_ac")
        , stringr::str_ends(x,"_kg_per_ha_lbsac") ~ stringr::str_replace(x,"_kg_per_ha_lbsac","_lbs_per_ac")
        , stringr::str_ends(x,"_per_ha_ac") ~ stringr::str_replace(x,"_per_ha_ac","_per_ac")
        , stringr::str_ends(x,"_area_ha_ac") ~ stringr::str_replace(x,"_area_ha_ac","_area_ac")
        , stringr::str_ends(x,"_kg_per_m2_lbsft2") ~ stringr::str_replace(x,"_kg_per_m2_lbsft2","_lbs_per_ft2")
        , stringr::str_ends(x,"_kg_per_m3_lbsft3") ~ stringr::str_replace(x,"_kg_per_m3_lbsft3","_lbs_per_ft3")
        , stringr::str_ends(x,"_m2_ft2") ~ stringr::str_replace(x,"_m2_ft2","_ft2")
        , TRUE ~ x
      )}
    )
}

###################################################################################
### stand-level summaries
###################################################################################
calc_silv_metrics <- function(tree_list, stand_area_ha = NULL, study_boundary = NULL, calc_imperial_units = F) {
  # get study area
  if(!is.null(study_boundary)){
    # bounds check
    if(
      !inherits(study_boundary,"sf")
      && !inherits(study_boundary,"sfc")
    ){stop("study_boundary must be sf class object")}
    if(is.na(sf::st_crs(study_boundary))){stop("study_boundary does not have a CRS")}
    if(inherits(study_boundary,"sf") && nrow(study_boundary)!=1){
      stop("study_boundary must only have a single record geometry")
    }
    if(inherits(study_boundary,"sfc") && length(study_boundary)!=1){
      stop("study_boundary must only have a single record geometry")
    }
    if(
      !all( sf::st_is(study_boundary, c("POLYGON","MULTIPOLYGON")) )
    ){
      stop("study_boundary must contain POLYGON type geometry only")
    }
    # area
    xxstand_area_ha <- study_boundary %>% 
      sf::st_area() %>% 
      as.numeric() %>% 
      `/`(10000)
  }else if(is.numeric(stand_area_ha)){
    xxstand_area_ha <- stand_area_ha[1]
  }else{
    stop("must provide `stand_area_ha` as numeric or `study_boundary` as sf object")
  }
  
  if(is.null(xxstand_area_ha) || is.na(xxstand_area_ha) || dplyr::coalesce(xxstand_area_ha,0)<=0){
    stop("could not determine valid stand_area_ha")
  }
  
  # summarize tree list
  if(!inherits(tree_list,"data.frame")){stop("tree_list must be data.frame class object")}
  
  ###### !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!! NEED TO PUT IN CHECKS FOR COLUMNS USED
  # see: cloud2trees::check_df_cols_all_missing()
  if( !any(stringr::str_detect(names(tree_list), "landfire_crown_biomass_kg")) ){
    tree_list <- tree_list %>% dplyr::mutate(landfire_crown_biomass_kg=as.numeric(NA))
  }
  if( !any(stringr::str_detect(names(tree_list), "cruz_crown_biomass_kg")) ){
    tree_list <- tree_list %>% dplyr::mutate(cruz_crown_biomass_kg=as.numeric(NA))
  }
  if( !any(stringr::str_detect(names(tree_list), "landfire_tree_kg_per_m3")) ){
    tree_list <- tree_list %>% dplyr::mutate(landfire_tree_kg_per_m3=as.numeric(NA))
  }
  if( !any(stringr::str_detect(names(tree_list), "cruz_tree_kg_per_m3")) ){
    tree_list <- tree_list %>% dplyr::mutate(cruz_tree_kg_per_m3=as.numeric(NA))
  }
  
  # agg
  agg <- tree_list %>%
    sf::st_drop_geometry() %>% 
    dplyr::ungroup() %>%
    dplyr::summarise(
      n_trees = dplyr::n()
      , mean_dbh_cm = mean(dbh_cm, na.rm = T)
      , mean_tree_height_m = mean(tree_height_m, na.rm = T)
      , mean_tree_cbh_m = mean(tree_cbh_m, na.rm = T)
      , loreys_height_m = sum(basal_area_m2*tree_height_m, na.rm = T) / sum(basal_area_m2, na.rm = T)
      , basal_area_m2 = sum(basal_area_m2, na.rm = T)
      , sum_dbh_cm_sq = sum(dbh_cm^2, na.rm = T)
      , landfire_crown_biomass_kg = sum(landfire_crown_biomass_kg, na.rm = F)
      , cruz_crown_biomass_kg = sum(cruz_crown_biomass_kg, na.rm = F)
      , mean_landfire_tree_kg_per_m3 = mean(landfire_tree_kg_per_m3, na.rm = T)
      , mean_cruz_tree_kg_per_m3 = mean(cruz_tree_kg_per_m3, na.rm = T)
    ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      stand_area_ha = xxstand_area_ha
      , trees_per_ha = (n_trees/stand_area_ha)
      , basal_area_m2_per_ha = (basal_area_m2/stand_area_ha)
      , qmd_cm = sqrt(sum_dbh_cm_sq/n_trees)
      , landfire_cfl_kg_per_m2 = landfire_crown_biomass_kg/(stand_area_ha*10000)
      , cruz_cfl_kg_per_m2 = cruz_crown_biomass_kg/(stand_area_ha*10000)
    ) %>%
    dplyr::select(-c(sum_dbh_cm_sq,landfire_crown_biomass_kg,cruz_crown_biomass_kg))
  # imperial
  if(calc_imperial_units){
    agg <- calc_imperial_units_fn(agg)
  }
  return(agg)
}
```

check out the data structure returned by the `calc_silv_metrics()` function

```{r}
calc_silv_metrics(tree_list = ans_temp$tree_list ,study_boundary = ans_temp$aoi, calc_imperial_units = T) %>% 
  dplyr::glimpse()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

### Study Area Summaries

let's calculate the metrics for all study areas

```{r, results=F, message=F}
# map the process over all data
silv_metrics_by_data <- 1:nrow(study_sites_processing_sf) %>% 
  purrr::map(function(x){
    # clip the data
    ans_temp <- cloud2trees::cloud2trees_to_lanl_trees(
      input_dir = study_sites_processing_sf$fdir[x]
      , study_boundary = study_sites_processing_sf %>% dplyr::slice(x)
      , bbox_aoi = F, buffer = 0 # this is the key to get only trees in our AOI
    )
    # silv metrics
    calc_silv_metrics(
      tree_list = ans_temp$tree_list 
      , study_boundary = ans_temp$aoi
      , calc_imperial_units = T
    ) %>% 
    dplyr::mutate(
      study_site = study_sites_processing_sf$study_site[x]
      , study_site_lab = study_sites_processing_sf$study_site_lab[x]
      , data_desc = study_sites_processing_sf$data_desc[x]
      , data_type = study_sites_processing_sf$data_type[x]
    )
  }) %>% 
dplyr::bind_rows()
# huh?
silv_metrics_by_data %>% dplyr::glimpse()
```

```{r, include=FALSE, eval=FALSE}
silv_metrics_by_data %>% 
dplyr::select(
  study_site,study_site_lab,data_desc,data_type
  , dplyr::all_of(c(
    "stand_area_ac"
    , "n_trees"
    , "trees_per_ac"
    , "mean_tree_height_ft"
    , "loreys_height_ft"
    , "mean_dbh_in"
    , "qmd_in"
    , "basal_area_ft2_per_ac"
  ))
) %>% 
tidyr::pivot_longer(
  cols = -c(study_site,study_site_lab,data_desc,data_type), names_to = "metric", values_to = "val"
) %>% 
# order metrics
dplyr::mutate(
  metric = factor(
    metric
    , ordered = T
    , levels = c(
        "stand_area_ac"
        , "n_trees"
        , "trees_per_ac"
        , "mean_tree_height_ft"
        , "loreys_height_ft"
        , "mean_dbh_in"
        , "qmd_in"
        , "basal_area_ft2_per_ac"
      )
    , labels = c(
        latex2exp::TeX("Stand Area (ac)", output = "character")
        , latex2exp::TeX("Number of Trees", output = "character")
        , latex2exp::TeX("Trees $ac^{-1}$")
        , latex2exp::TeX("Mean Tree Height (ft)", output = "character")
        , latex2exp::TeX("Lorey's Mean Height (ft)", output = "character")
        , latex2exp::TeX("Mean DBH (in)", output = "character")
        , latex2exp::TeX("QMD (in)", output = "character")
        , latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
      )
  )
) %>% 
# dplyr::mutate(val = scales::comma(val,accuracy=0.1)) %>% 
ggplot2::ggplot(
  mapping = ggplot2::aes(y = study_site_lab, x = val, color = data_type)
) +
ggplot2::geom_point() +
ggplot2::facet_wrap(facets = dplyr::vars(metric), scales = "free_x", labeller = ggplot2::label_parsed) +
ggplot2::theme_light()
```

#### Metric units

```{r}
silv_metrics_by_data %>% 
  dplyr::mutate(
    nlab = paste0(study_site_lab,"<br>",scales::comma(stand_area_ha,accuracy=1)," ha")
  ) %>% 
  dplyr::relocate(nlab) %>% 
  dplyr::select(
    nlab
    , data_desc
    # different
    , n_trees
    , trees_per_ha
    , mean_tree_height_m
    , loreys_height_m
    , mean_dbh_cm
    , qmd_cm
    , basal_area_m2_per_ha
    , mean_landfire_tree_kg_per_m3
    , mean_cruz_tree_kg_per_m3
) %>% 
dplyr::mutate(
  dplyr::across(c(n_trees), ~scales::comma(.x,accuracy=1))
  , dplyr::across(
    -c(nlab,data_desc,n_trees,tidyselect::ends_with("3"))
    , ~scales::comma(.x,accuracy=0.1)
  )
  , dplyr::across(
    tidyselect::ends_with("3")
    , ~scales::comma(.x,accuracy=0.001)
  )
) %>% 
kableExtra::kbl(
  caption = "Stand summary metrics in metric units"
  , col.names = c(
    "site","collection"
    , "trees"
    , "trees ha<sup>-1</sup>"
    , "mean<br>tree ht. (m)"
    , "Lorey's<br>tree ht. (m)"
    , "mean<br>DBH (cm)"
    , "QMD (cm)"
    , "BA (m<sup>2</sup> ha<sup>-1</sup>)"
    , "mean<br>CBD (kg m<sup>-3</sup>)<br>LANDFIRE"
    , "mean<br>CBD (kg m<sup>-3</sup>)<br>Cruz"
  )
  , escape = F
) %>% 
kableExtra::kable_styling(font_size = 12) %>% 
kableExtra::collapse_rows(columns = 1:2, valign = "top")  

```

#### Imperial units

```{r}
silv_metrics_by_data %>% 
  dplyr::mutate(
    nlab = paste0(study_site_lab,"<br>",scales::comma(stand_area_ac,accuracy=1)," ac")
  ) %>% 
  dplyr::relocate(nlab) %>% 
  dplyr::select(
    nlab
    , data_desc
    # different
    , n_trees
    , trees_per_ac
    , mean_tree_height_ft
    , loreys_height_ft
    , mean_dbh_in
    , qmd_in
    , basal_area_ft2_per_ac
    , mean_landfire_tree_lbs_per_ft3
    , mean_cruz_tree_lbs_per_ft3
) %>% 
dplyr::mutate(
  dplyr::across(c(n_trees), ~scales::comma(.x,accuracy=1))
  , dplyr::across(
    -c(nlab,data_desc,n_trees,tidyselect::ends_with("3"))
    , ~scales::comma(.x,accuracy=0.1)
  )
  , dplyr::across(
    tidyselect::ends_with("3")
    , ~scales::comma(.x,accuracy=0.001)
  )
) %>% 
kableExtra::kbl(
  caption = "Stand summary metrics in imperial units"
  , col.names = c(
    "site","collection"
    , "trees"
    , "trees ac<sup>-1</sup>"
    , "mean<br>tree ht. (ft)"
    , "Lorey's<br>tree ht. (ft)"
    , "mean<br>DBH (in)"
    , "QMD (in)"
    , "BA (ft<sup>2</sup> ac<sup>-1</sup>)"
    , "mean<br>CBD (lb ft<sup>-3</sup>)<br>LANDFIRE"
    , "mean<br>CBD (lb ft<sup>-3</sup>)<br>Cruz"
  )
  , escape = F
) %>% 
kableExtra::kable_styling(font_size = 12) %>% 
kableExtra::collapse_rows(columns = 1:2, valign = "top")  
```

