# Point Cloud Processing Results

## Processing Time

let's look into the processing time comparison across data sets

the `cloud2trees::cloud2trees()` program automatically tracks processing time for all sections and stores the tracking data in the "point_cloud_processing_delivery" directory in a file called "processed_tracking_data.csv"

```{r}
dir_temp <- "../data"
# what processing data?
df_temp <-
  list.files(
    dir_temp
    , pattern = "processed_tracking_data.csv"
    , recursive = T
  ) %>% 
  dplyr::tibble() %>% 
  setNames("fpath") %>% 
  dplyr::mutate(
    data_desc = stringr::word(fpath, 2, sep = "/") %>% 
      stringr::str_remove_all("_processing") %>% 
      stringr::str_replace_all("_"," ") %>% 
      stringr::str_squish() %>% 
      toupper()
    , data_type = data_desc %>% 
      stringr::str_remove_all("[0-9]") %>% 
      stringr::str_squish()
    , study_site = stringr::word(fpath, sep = "/")
    , fpath = file.path(dir_temp, fpath)
    , fdir = dirname(fpath)
  )
# read in processing data
df_temp <- 1:nrow(df_temp) %>% 
  purrr::map(\(x)
    readr::read_csv(
      df_temp$fpath[x]
      , show_col_types = F
      , progress = F
    ) %>%
    dplyr::mutate(
      study_site = df_temp$study_site[x]
      , data_desc = df_temp$data_desc[x]
    )
  ) %>% 
  dplyr::bind_rows() %>% 
  dplyr::inner_join(df_temp, by = dplyr::join_by(study_site,data_desc))
# add in study bounds
study_sites_processing_sf <- 
  study_sites_sf %>% 
  dplyr::select(study_site, study_site_lab) %>% 
  dplyr::inner_join(
    df_temp
    , by = "study_site"
    , relationship = "one-to-many"
  )
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

let's relativize and proportionalize the tracking data

```{r}
# aggregate the total processing time
study_sites_processing_sf <- 
  study_sites_processing_sf %>% 
  dplyr::mutate(
    timer_total_time_mins = timer_cloud2raster_mins + timer_raster2trees_mins +
      timer_trees_dbh_mins + timer_trees_cbh_mins + timer_trees_type_mins +
      timer_trees_hmd_mins + timer_trees_biomass_mins + timer_write_data_mins
    , timer_tree_extraction_mins = timer_cloud2raster_mins + timer_raster2trees_mins
    , las_area_ha = (las_area_m2/10000)
    , points_m2 = number_of_points/las_area_m2
    # relative
    , dplyr::across(
      .cols = tidyselect::starts_with("timer_") & tidyselect::ends_with("_mins")
      , .fns = ~ (.x*60)/las_area_ha # sec/ha
      # , .fns = ~ .x/las_area_ha # min/ha
      , .names = "{.col}_secperha"
      # , .names = "{.col}_minperha"
    )
    # proportion
    , dplyr::across(
       .cols = c(timer_tree_extraction_mins,
        timer_trees_dbh_mins, timer_trees_cbh_mins, timer_trees_type_mins,
        timer_trees_hmd_mins, timer_trees_biomass_mins, timer_write_data_mins)
       , .fns = ~ .x/timer_total_time_mins
       , .names = "{.col}_pct"
    )
  )
study_sites_processing_sf %>% dplyr::glimpse()
```

that's a lot of tracking data, let's look at the main timing parameters

```{r}
# format data for plotting/tabling
table_temp <-
  study_sites_processing_sf %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(
    study_site, study_site_lab, data_desc, data_type, number_of_points, las_area_ha, points_m2
    , c(timer_tree_extraction_mins,
      timer_trees_dbh_mins, timer_trees_cbh_mins, timer_trees_type_mins,
      timer_trees_hmd_mins, timer_trees_biomass_mins, timer_write_data_mins
      , timer_total_time_mins
      , c(tidyselect::ends_with("_pct") & tidyselect::starts_with("timer_"))
      , c(tidyselect::ends_with("_secperha") & tidyselect::starts_with("timer_"))
    )
  ) %>% 
  tidyr::pivot_longer(
    cols = -c(study_site, study_site_lab, data_desc, data_type, number_of_points, las_area_ha, points_m2)
  ) %>% 
  dplyr::mutate(
    units = stringr::word(name, -1, sep = "_")
    , section = name %>% 
      stringr::str_remove_all("timer_") %>% 
      stringr::str_remove_all("_mins") %>% 
      stringr::str_remove_all("_pct") %>% 
      stringr::str_remove_all("_secperha")
  ) %>% 
  dplyr::select(-name) %>% 
  # dplyr::count(units)
  tidyr::pivot_wider(names_from = units, values_from = value) %>% 
  dplyr::mutate(
    mins_lab = scales::comma(mins,accuracy = 0.1)
    , perha_lab = scales::comma(secperha,accuracy = 0.01)
    , pct_lab = scales::percent(pct,accuracy = 0.1)
    # site lab
    , big_lab = stringr::str_c(
      data_desc
      , paste0("area: ", scales::comma(las_area_ha, accuracy = 1, suffix = " ha"))
      , paste0("points: ", scales::comma(number_of_points, accuracy = 0.1, scale = 1/1000000, suffix = "M"))
      , paste0("points m<sup>-2</sup>: ", scales::comma(points_m2, accuracy = 0.1))
      , sep = "<br>"
    )
    , big_lab_ggplot = stringr::str_c(
      data_desc
      , paste0("area: ", scales::comma(las_area_ha, accuracy = 1, suffix = " ha"))
      , paste0("points: ", scales::comma(number_of_points, accuracy = 0.1, scale = 1/1000000, suffix = "M"))
      , paste0("points/m2: ", scales::comma(points_m2, accuracy = 0.1))
      , sep = "\n"
    )
  ) %>%  
  # dplyr::count(section)
  dplyr::filter(!is.na(mins)) %>% 
  # section lab after filter so factor doesn't have rand levels
  dplyr::mutate(
    section = section %>% 
      stringr::str_replace_all("_", " ") %>% 
      stringr::str_remove_all("time") %>% 
      stringr::str_replace_all("dbh", "DBH") %>% 
      stringr::str_replace_all("cbh", "CBH") %>% 
      stringr::str_replace_all("hmd", "HMD") %>% 
      stringr::str_squish() %>% 
      forcats::fct_inorder()
  )
  
# table it
table_temp %>% 
  dplyr::select(study_site_lab,big_lab,section,mins_lab, perha_lab, pct_lab) %>% 
  kableExtra::kbl(
    caption = "Point cloud processing section run time"
    , col.names = c(
      "Site", "Data"
      , "Processing section"
      , "time (minutes)"
      , "seconds per ha"
      , "% of total time"
      )
    , escape = F
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::collapse_rows(columns = 1:2, valign = "top")

```

that's a lot of numbers to digest, let's plot the data

### Point Cloud Processing Time versus Point Density

```{r}
# per ha time based on point density
ggplot2::ggplot(
  data = table_temp %>% dplyr::filter(section == "total")
  , mapping = ggplot2::aes(y = secperha, x = points_m2, color = data_type)
) +
  ggplot2::geom_point(size = 4, alpha = 0.9) +
  ggplot2::scale_color_viridis_d(option = "magma", begin = 0.1, end = 0.5) +
  ggplot2::labs(
    x = latex2exp::TeX("points $m^{-2}$")
    , y = latex2exp::TeX("total seconds $ha^{-1}$")
    , color = ""
    , subtitle = "Point Cloud Processing Time versus Point Density"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "top")
```

### Point Cloud Processing Time by Section

```{r}
table_temp %>% 
  dplyr::filter(section != "total") %>% 
  dplyr::mutate(section = forcats::fct_rev(section)) %>% 
ggplot2::ggplot(
  mapping = ggplot2::aes(y = big_lab_ggplot, x = pct, fill = section, group = section)
) +
  ggplot2::geom_col(
    width = 0.7, alpha=0.8
  ) +
  ggplot2::geom_text(
    mapping = ggplot2::aes(
        label = scales::percent(ifelse(pct>=0.06,pct,NA), accuracy = 1)
        , fontface = "bold"
      )
    , position = ggplot2::position_stack(vjust = 0.5)
    , color = "black", size = 4
  ) +
  ggplot2::facet_wrap(facets = dplyr::vars(study_site_lab), scales = "free_y") + 
  ggplot2::scale_fill_viridis_d(option = "turbo", begin = 0.1, end = 0.9) +
  ggplot2::scale_x_continuous(labels = scales::percent_format()) +
  labs(
    fill = "", y = ""
    , x = "% Point Cloud Total Processing Time"
    , subtitle = "Point Cloud Processing Time by Section"
  ) +
  theme_light() +
  theme(
    legend.position = "top"
    , legend.direction  = "horizontal"
    , legend.title = element_text(size=7)
    , axis.title.x = element_text(size=10, face = "bold")
    , axis.title.y = element_text(size = 8)
    , axis.text.x = element_blank()
    , axis.text.y = element_text(color = "black",size=10, face = "bold")
    , axis.ticks.x = element_blank()
  ) +
  guides(
    fill = guide_legend(nrow = 3, byrow = T, reverse = T, override.aes = list(alpha = 0.9))
  )
```

