# `cloud2trees` to LANL TREES{#lanl_trees}

we format the `cloud2trees::cloud2trees()` outputs to meet the requirements of the [LANL TREES](https://github.com/lanl/Trees/) program, which prepares data for fire modeling. This data formatting process is accomplished via `cloud2trees::cloud2trees_to_lanl_trees()`. The current iteration of this function requires the user to specify surface fuel load parameters, such as litter and herbaceous/grass fuel loads, which are assumed constant across the study area. 

## Surface Fuel Quantification

let's pull in surface fuel loading parameters which were determined through a literature review ([Gallagher et al. 2017](https://www.fs.usda.gov/rds/archive/catalog/RDS-2017-0061); [Ottmar & Andreu 2007](https://iris.fws.gov/APPS/ServCat/DownloadFile/25710))

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

```{r}
surface_fuels <-
  readxl::read_excel("../data/Surface_Fuels.xlsx") %>% 
  dplyr::mutate(dplyr::across(
    .cols = tidyselect::contains("%")
    , ~ .x*.01
  )) %>% 
  dplyr::rename_with(
    ~ stringr::str_replace_all(.x,"%","pct") %>% 
      stringr::str_replace_all("[ .]", "_") %>% 
      stringr::str_replace_all("[^[:alnum:]_]+", "") %>% 
      tolower()
  ) %>% 
  dplyr::rename(study_site=site) %>% 
  dplyr::mutate(
    study_site = study_site %>% stringr::str_to_title() %>% stringr::str_remove_all("[ .]")
    # these are somewhat based on the literature assuming Fall timing
    , herbaceous_moisture_pct = dplyr::case_when(
      is.na(herbaceous_moisture_pct) & study_site == "SycanMarsh" ~ 0.3
      , is.na(herbaceous_moisture_pct) & study_site == "FortStewart" ~ 0.95
      , is.na(herbaceous_moisture_pct) & study_site == "SaltCabin" ~ 0.3
      , is.na(herbaceous_moisture_pct) & study_site == "NewJersey" ~ 0.6
      , T ~ herbaceous_moisture_pct
    )
  ) %>% 
  # put on our fancy labels
  dplyr::inner_join(
    study_sites_sf %>% sf::st_drop_geometry() %>% dplyr::select(study_site, study_site_lab)
    , by = "study_site"
  )
# add it to our processing data table
study_sites_processing_sf <- study_sites_processing_sf %>% 
  dplyr::left_join(surface_fuels %>% dplyr::select(-study_site_lab),by="study_site")

# table it
surface_fuels %>% 
  tidyr::pivot_longer(
    cols = -tidyselect::starts_with("study_site")
  ) %>% 
  dplyr::mutate(
    fuel_type = stringr::word(name,sep="_")
    , name = stringr::str_remove(name, paste0(fuel_type,"_"))
  ) %>%
  tidyr::pivot_wider(names_from = name, values_from = value) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::ends_with("_pct"), ~scales::percent(.x,accuracy=1)
    )
    # , dplyr::across(
    #   dplyr::where(is.numeric), ~scales::comma(.x)
    # )
  ) %>% 
  dplyr::select(-study_site) %>% 
   kableExtra::kbl(
    caption = "surface fuel loading parameters determined through a literature review"
    , col.names = c(
      "site","layer"
      , "Bulk Density<br>(kg m<sup>-3</sup>)"
      , "Height (m)"
      , "Moisture"
      , "Surface Area<br>to Volume (m)"
    )
    , escape = F
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::collapse_rows(columns = 1:2, valign = "top")  
```

## `cloud2trees::cloud2trees_to_lanl_trees()` example process

let's use Sycan Marsh (OR) to show how easy it is to use `cloud2trees::cloud2trees_to_lanl_trees()`

```{r, message=FALSE, results=F}
# just one record with surface fuels for ease
processing_info_temp <- study_sites_processing_sf %>% 
  dplyr::filter(
    study_site=="SycanMarsh"
    & data_type=="ALS"
  )
# save a dir path
my_dir_temp <- tempdir()
# pass info to the fn
cloud2trees_to_lanl_trees_ans_temp <- cloud2trees::cloud2trees_to_lanl_trees(
  input_dir = processing_info_temp$fdir # point_cloud_processing_delivery 
  , study_boundary = processing_info_temp # sf data
  , bbox_aoi = T # use a bbox
  , buffer = 20
  , topofile = "flat"
  , cbd_method = "landfire"
  , output_dir = my_dir_temp
  , fuel_litter = list(
    ilitter = 1
    , lrho = processing_info_temp$litter_bulk_density_kg_m3
    , lmoisture = processing_info_temp$litter_moisture_pct
    , lss = processing_info_temp$litter_surface_area_to_volume_m
    , ldepth = processing_info_temp$litter_height_m
  )
  , fuel_grass = list(
    igrass = 1
    , grho = processing_info_temp$herbaceous_bulk_density_kg_m3
    , gmoisture = processing_info_temp$herbaceous_moisture_pct
    , gss = processing_info_temp$herbaceous_surface_area_to_volume_m
    , gdepth = processing_info_temp$herbaceous_height_m
  )
)
```

let's see what we got

```{r}
cloud2trees_to_lanl_trees_ans_temp %>% names()
```

check out the tree list points that were cropped to our study boundary with a bounding box and a 20m buffer

```{r}
ggplot2::ggplot() +
  ggplot2::geom_sf(
    data = cloud2trees_to_lanl_trees_ans_temp$aoi
    , fill = NA, color = "navy", lwd = 2
  ) +
  ggplot2::geom_sf(
    data = cloud2trees_to_lanl_trees_ans_temp$tree_list
    , mapping = ggplot2::aes(color = tree_height_m)
  ) +
  ggplot2::scale_color_distiller(palette = "Blues", direction = 1) +
  ggplot2::labs(color = "Height (m)") +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "top")
```

check out the DTM that was set to 2m resolution by default and cropped to our study boundary with a bounding box and a 20m buffer

```{r}
terra::plot(cloud2trees_to_lanl_trees_ans_temp$dtm, main = "DTM (m)")
terra::plot(
  cloud2trees_to_lanl_trees_ans_temp$aoi %>%
    sf::st_transform(terra::crs(cloud2trees_to_lanl_trees_ans_temp$dtm)) %>%
    terra::vect()
  , add = T, border = "navy", col = NA
  , lwd = 11
)
```

and what are all of these file paths?

```{r}
# the path to the "Lidar_Bounds.geojson" file
cloud2trees_to_lanl_trees_ans_temp$domain_path  
# the path to the "topo.dat" file
cloud2trees_to_lanl_trees_ans_temp$topofile_path  
# the path to the TREES program "fuellist" file
cloud2trees_to_lanl_trees_ans_temp$fuellist_path  
# the path to the "Cloud2Trees_TreeList.txt" file
cloud2trees_to_lanl_trees_ans_temp$treelist_path  
```

the `cloud2trees::cloud2trees_to_lanl_trees()` program made a delivery directory for the astute readers

```{r}
list.files(file.path(my_dir_temp,"lanl_trees_delivery"))
```

let's look at a few records of the fuellist

```{r}
readLines(cloud2trees_to_lanl_trees_ans_temp$fuellist_path)[1:11]
```

let's look at the structure of the "Cloud2Trees_TreeList.txt" file which does not include column names

```{r}
readr::read_delim(
  cloud2trees_to_lanl_trees_ans_temp$treelist_path
  , col_names = F
  , progress = F, show_col_types = F
) %>% 
dplyr::glimpse()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## `cloud2trees::cloud2trees_to_lanl_trees()` everything

as a reminder we processed all of these data

```{r}
study_sites_processing_sf %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(study_site_lab, data_desc, number_of_points, las_area_ha, fdir) %>% 
  dplyr::mutate(dplyr::across(dplyr::where(is.numeric),~scales::comma(.x))) %>% 
  dplyr::rename(cloud2trees_output_dir = fdir) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling(font_size = 12)
```

`cloud2trees::cloud2trees_to_lanl_trees()` all of the `cloud2trees::cloud2trees()` processed data

```{r, results=F, message=FALSE, warning=FALSE}
# dir
dir_temp <- file.path("c:/Users/georg/Downloads/", "cloud2trees_TREES_data")
if(!dir.exists(dir_temp)){
  dir.create(dir_temp)
}
# start at timer
st_temp <- Sys.time()
# every data
ans_temp <-
  1:nrow(study_sites_processing_sf) %>%
  purrr::map(function(x){
    # delivery folder
    fp <- file.path(
      dir_temp
      , paste0(
        study_sites_processing_sf$study_site[x]
        , "_", stringr::str_replace_all(study_sites_processing_sf$data_desc[x],"[ .]", "_")
      )
    )
    if(!dir.exists(fp)){dir.create(fp)}
    
    # cloud2trees::cloud2trees_to_lanl_trees()
    cloud2trees::cloud2trees_to_lanl_trees(
      input_dir = study_sites_processing_sf$fdir[x] # point_cloud_processing_delivery 
      , study_boundary = study_sites_processing_sf %>% dplyr::slice(x) # sf data
      , bbox_aoi = T # use a bbox
      , buffer = 30
      , topofile = "flat"
      , cbd_method = "landfire"
      , output_dir = fp
      , fuel_litter = list(
        ilitter = 1
        , lrho = study_sites_processing_sf$litter_bulk_density_kg_m3[x]
        , lmoisture = study_sites_processing_sf$litter_moisture_pct[x]
        , lss = study_sites_processing_sf$litter_surface_area_to_volume_m[x]
        , ldepth = study_sites_processing_sf$litter_height_m[x]
      )
      , fuel_grass = list(
        igrass = 1
        , grho = study_sites_processing_sf$herbaceous_bulk_density_kg_m3[x]
        , gmoisture = study_sites_processing_sf$herbaceous_moisture_pct[x]
        , gss = study_sites_processing_sf$herbaceous_surface_area_to_volume_m[x]
        , gdepth = study_sites_processing_sf$herbaceous_height_m[x]
      )
    )
  })
# end a timer
end_temp <- Sys.time()
secs_temp <- difftime(end_temp,st_temp,units="secs") %>% as.numeric()
# set names
names(ans_temp) <- paste0(
  study_sites_processing_sf$study_site
  , "_", stringr::str_replace_all(study_sites_processing_sf$data_desc,"[ .]", "_")
)
```

to process all of that data took `r scales::comma(secs_temp,accuracy=0.1)` seconds (`r scales::comma(secs_temp/60,accuracy=0.01)` minutes), not bad, not bad

just to confirm we got the data we expected, let's look at one result using the Sycan Marsh UAS Lidar 2023 data 

```{r}
cloud2trees_to_lanl_trees_ans_temp <- ans_temp$SycanMarsh_UAS_LIDAR_2023
cloud2trees_to_lanl_trees_ans_temp %>% names()
```

check out the tree list points that were cropped to our study boundary with a bounding box and a 20m buffer

```{r}
ggplot2::ggplot() +
  ggplot2::geom_sf(
    data = cloud2trees_to_lanl_trees_ans_temp$aoi
    , fill = NA, color = "navy", lwd = 2
  ) +
  ggplot2::geom_sf(
    data = cloud2trees_to_lanl_trees_ans_temp$tree_list
    , mapping = ggplot2::aes(color = tree_height_m)
  ) +
  ggplot2::scale_color_distiller(palette = "Blues", direction = 1) +
  ggplot2::labs(color = "Height (m)") +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "top")
```

check out the DTM that was set to 2m resolution by default and cropped to our study boundary with a bounding box and a 20m buffer

```{r}
terra::plot(cloud2trees_to_lanl_trees_ans_temp$dtm, main = "DTM (m)")
terra::plot(
  cloud2trees_to_lanl_trees_ans_temp$aoi %>%
    sf::st_transform(terra::crs(cloud2trees_to_lanl_trees_ans_temp$dtm)) %>%
    terra::vect()
  , add = T, border = "navy", col = NA
  , lwd = 11
)
```

and what are all of these file paths?

```{r}
# the path to the "Lidar_Bounds.geojson" file
cloud2trees_to_lanl_trees_ans_temp$domain_path  
# the path to the "topo.dat" file
cloud2trees_to_lanl_trees_ans_temp$topofile_path  
# the path to the TREES program "fuellist" file
cloud2trees_to_lanl_trees_ans_temp$fuellist_path  
# the path to the "Cloud2Trees_TreeList.txt" file
cloud2trees_to_lanl_trees_ans_temp$treelist_path  
```

let's look at a few records of the fuellist

```{r}
readLines(cloud2trees_to_lanl_trees_ans_temp$fuellist_path)[1:11]
```

let's look at the structure of the "Cloud2Trees_TreeList.txt" file which does not include column names

```{r}
readr::read_delim(
  cloud2trees_to_lanl_trees_ans_temp$treelist_path
  , col_names = F
  , progress = F, show_col_types = F
) %>% 
dplyr::glimpse()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

